---
title: "Tibbles vs. data.frame"
author: "Lauren Kapraun"
date: "09/17/18"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
install.packages("tibble")
library(tibble)
library(tidyverse)
library(tidyr)
library(magrittr)
```

Before starting, I installed the following: **tibble** packages and the **tibble, tidyverse, tidyr, magrittr** libraries. 

The two main fifferences in using a tibble versus the `data.frame` is printing and subsetting.

## Printing
When using tibbles, the refined print method only displays the first 10 rows by default.

```{r defaultPrint, message=FALSE, warning=FALSE}
tb = tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
```

### Display Options
One option you can use to display more than the first 10 rows is to use the `print(n)` function, where "n" is the number of rows you want to display. If some of the columns are hidden due to to size, you can also use `width=Inf` to show all of the columns in the table. 

```{r printFifteenRows}
tb %>%
  print(n = 15, width = Inf)
```

#### Display Min & Max
To show a minimum or a maximum number of rows use: `options(tibble.print_max = n, tribble.print_min = m)`. In this instance, if there are more than "m" rows, it will display all the rows up to "n" rows. 

```{r printMinMax, results='hide'}
options(tibble.print_max = 30, tibble.print_min = 20)
  tb
```

#### Disply All Rows
To always display all rows, use: `options(dplyr.print_min = Inf)`.

```{r printAllRows, results='hide'}
options(dplyr.print_min = Inf)
  tb
```

#### Disply All Columns
To show all columns by default, use: `options(tribble.width = Inf)`

```{r printAllColumns, results='hide'}
options(n = 15, tribble.width = Inf)
  tb
```

#### Disply All in New Tab
You can also use RStudio's built-in data viewer to display a scrollable view of the dataset using `View()`. 

### Final Print Options
To view all of the package options, run `package?tibble`

## Subsetting
To pull from a single vairable, you can use `[[` to extract by name or position, or you can use `$` to extract only by name.

Tibbles are more strict than data.frames, as they do not allow for partial matching.

```{r setupSubset}
df = tibble(
  x = runif(5),
  y = runif(5)
)

# Extract by Name
df$x
df[["x"]]

# Extract by Position
df[[1]]

# To Use in a Pipe, Use a Special Placeholder
df %>% .$x
df %>% .[["x"]]
```

## Interacting with Older Code
If you are trying to use an older function that doesn't seem to be working, use `as.data.frame()` to convert the tibble to a data.frame.

```{r switchToDataFrame, results='hide'}
class(as.data.frame(tb))
```